Supp mat: time series paper

### 2017-05-11 writing/supp_mat_draft.tex
made changes to formatting and introduction

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=0.75in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{float} \usepackage{setspace} \usepackage{lineno}
\usepackage[round]{natbib} \bibpunct[; ]{(}{)}{,}{a}{}{;} \linenumbers

@@

\begin{spacing}{1.5} %use 2
    \begin{flushleft}
Running head: 
\vspace{3 mm}

Title: Your time series is (probably) too short
\vspace{7 mm}

Author(s): \textsc{Easton R. White$^{1,2}$} 
\vspace{3 mm}

Addresses: \\ \emph{$^1$ Center for Population Biology, University of California-Davis, Davis, California 95616 USA}
\vspace{3 mm}

$^2$Corresponding author: eawhite@ucdavis.edu
\vspace{3 mm}

Number of words: X,XXX 
\vspace{3 mm}

To be submitted to: \emph{XXXXX} (Article)
\vspace{3 mm}

Keywords:

\vspace{3 mm}

\end{flushleft}
\end{spacing}

\linenumbers

\clearpage

\subsection{Abstract}\label{abstract}

Long-term census data is necessary to better understand ecological
processes and make management decisions. In addition, census data is
often expensive, requiring a lot of people-hours or equipment. However,
little work has actually addressed the length of time series required.
In other words, when is a time series of census data long enough to
address a question of interest? Here, I explore two approaches to
address this question: one simulation-based and an empirical approach. I
specifically determine the minimum time series length required to
estimate significant increases or decreases in population abundance.
Importantly, I examine the ability to detect a trend with a set level of
statistical power, which is often neglected in ecological time series
analyses. Using simple simulations, I demonstrate how the minimum time
series length required increases with weaker trends in abundance and
with higher variability in population size. In addition, I examine 868
populations of vertebrate species to determine the minimum time required
to detect changes in their abundance. Here I show 10-15 years of
continuous monitoring are required in order to achieve a high level of
statistical power. Similar to simulation-approaches, the minimum time
required for field census data strongly depends on trend strength,
population variability, and temporal autocorrelation. Perhaps
surprisingly, the minimum time required did not correlate well with
biological explanatory variables, like body size or generation length.
These results point to the importance of sampling populations over long
periods of time. Here, I stress the need for verifying a study has
enough statistical power to accurately determine long-term changes in
population abundance. Most studies, especially those less than 10-15
years, are probably under-powered and potentially misleading.

\subsection{Introduction}\label{introduction}

\begin{itemize}
\tightlist
\item
  intro long time series
\item
  my general question
\item
  primer on power
\item
  more general power interpretation
\item
  2+ approaches
\item
  what we do in this paper
\end{itemize}

Observational studies, and censusing populations, has become a
cornerstone of modern ecology and conservation biology (Magurran et al.
2010). Longterm data is necessary to understand population dynamics and
to assess extinction risk of species. Even though many time series may
now be considered ``longterm'', most are still short (i.e.~less than X
years). \emph{I could have statistic on how long is considered
``longterm'' in papers}. Time series are typically short for a variety
of reasons. They are often coupled with an experiment, which may only
last a couple of years. In addition, short funding cycles make it
difficult to examine populations over longer periods of time. There are
of course notable exceptions including the Isle Royale study, Iowa corn
study, cod, lynx-hair, and many others.

Naturally, one might ask, how long of a time series is actually
necessary? This question has important scientific and management
implications. Scientifically, we need to know how long a time series is
required to address a specific question. Too short of a time series may
lead to wrong conclusions given large natural year-to-year variability.
From a management perspective, it is important to understand when a
trend in abundance over time is actually meaningful or not. For instance
the International Union for the Conservation of Nature (IUCN)
suggest\ldots{}In addition, because sampling is typically expensive, we
do not want to over sample (Gerber, DeMaster, and Kareiva 1999).

In determining the number of samples required for a particular
experiment four quantities are intricately linked: statistical power,
effect size, alpha, sample size (Legg and Nagy 2006). The exact
relationship between these quantities depends on the specific
statistical test. Formally, statistical power is one minus the
probability of a type II error, or false negative. In other words, it is
a measure of how well you could detect a trend given it is actually
significant. Prior to an experiment, one could set appropriate levels of
power, \(\alpha\), and the effect size to estimate the sample size
required for the experiment. This approach, however, is not as simple
for a time series as the data points are clearly non-independent.

For time series data, two general approaches can be used. First, as can
be done with experimental data, we can relax the definition of
statistical power (cite Bolker). Instead of using classic statistical
tests (e.g.~t-test), and corresponding power calculations, simulation
approaches can be used (Gerrodette 1987). Simple population models can
be simulated with parameter values corresponding to a population of
interest. Out of a large number of simulations, power can be calculated
as the fraction of simulations that meet some criteria. The specific
criteria depend on the question at hand. Then one can run simulations
for time series of different lengths to determine the minimum time
series length required to be confident that you will be able to answer
your question. In addition to using simulated data, empirical time
series can also be used. We usually do not have multiple replicates of
similar populations, but we can subsample an empirical time series.
Subsamples of different lengths can then be evaluated to see which
fraction of subsamples meet some criteria, again a measure of
statistical power\ldots{}

Past work has investigated these questions with either theoretical
approaches not using data or by examining a single population. As an
example of the former, (Rhodes and Jonzen 2011) examined..

In this paper, we use both simulations and a database of empirical time
series to determine the minimum number of years required to address
various questions. To demonstrate these techniques, we address three
specific questions: (1) what is the minimum time series length required
to identify significant changes over time using linear regression, (2)

\subsection{Methods}\label{methods}

\subsubsection{Theoretical approach}\label{theoretical-approach}

One approach to determining the minimum time series needed is through
repetitive simulations of a population model (Gerrodette 1987). This is
the same approach one might use in sample size calculations for any
experimental design too complicated for simple power analyses (Johnson
et al. 2015). We only briefly discuss this approach as it has been
desribed elsewhere (citations). Essentially, we take a stochastic
population model and simulate it for a number of years repetitively. We
can then determine the minimum time series required to achieve designed
levels of significance and power. This approach requires us to determine
values for our model parameters. As an example, we can take the
following population model for population size \(N\) at time \(t\): \[
N(t + 1) = N(t) + r(t) \mbox{ with } r(t) \sim N(\phi \mu, \sigma)
\] where \(\epsilon\) is a normally-distributed random noise term with
mean \(\mu\) and standard deviation \(\sigma\). The rate of growth \(r\)
is also the trend strength of increase or decrease. Lastly, \(\phi\)
represents an autocorrelation term where \(\phi=0\) indicates no
autocorrelation, \(\phi>0\) is positively autocorrelated. This model
represents linear growth with autocorrelation and noise. However, it is
certainly possible to use any other population model.

By simulating this model repetitively for different lengths of time, we
can determine the minimum time series required. Here statistical power
is simply the number of trials that are themselves significant, as the
real trend is different from zero. Then the minimum time required is the
minimum time required to obtain statistical power above some
pre-determined threshold. Historically, 0.8\ldots{}.s

In figure X, a number of simulated time series are shown for a set
number of time periods (t=40). It is clear that statistical power
increases quickly with increases in \(t\) (Fig. Xb). Once power is
greater than XX (the dotted line), that is the minimum time required
(\(t_{min}\)

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{theoretical approach\label{fig:theoretical_approach}}
\end{figure}

\subsubsection{Data source}\label{data-source}

We use a database of 2444 population time series compiled in (Keith et
al. 2015). The data is originally from the Global Population Dynamics
Database (NERC Centre for Population Biology 2010). We filtered out
short time series (less than 35) and were left with 1019 time. The
database includes information on 499 vertebrate species with a focus on
mammals, birds, and fish. The database also includes information on
generation length and census specifications. For each time series, we
also calculated variables of interest like variance in population size,
longterm trend in abundance (slope coefficient from simple linear
regression), and temporal autocorrelation. Lastly, we also calculated
the minimum number of years required to sample according to IUCN
standards under rule \ldots{} (citation needed).

For a subset of populations in the database (X = XX), we were able to
connect a database on biological characteristics like body size and
generation length. We used

\subsubsection{Empirical approach}\label{empirical-approach}

Our objective is to determine the minimum time series length
(\(T_{min}\)) required to answer a specific question of interest. As we
describe below, the minimum time series length calculation is particular
to each question. Intuitively, we might expect a longer minimum time
series length required given more complicated questions.

We assume that each time series in our database are long enough to
include all necessary information (e.g.~variability) about the
population. In other words, they are a Representative sample. We first
take all possible contiguous subsamples of each time series. For
example, a time series of 40 years would have 39 possible contiguous
subsamples of length 2, 38 possible contiguous subsamples of length 3,
and so forth. In this paper, we will try to ask a simple, specific
question. We seek to determine the \(T_{min}\) required to be confident
in an estimate of the slope coefficient from linear regression.
Therefore, we run a linear regression for each subsampled time series.
Then, we determine the fraction of subsamples of a particular length
that have slopes from linear regression which are statistically
different from zero given the longterm, or ``true``, time series also
has a significant slope. This is a measure of statistical power. Lastly,
we determine which subsample length is required to guarantee we have
above a certain threshold of statistical power, which here we will
assume is 0.8 by convention (citation).

We also examine how minimum time series calculations changed depending
on the question of interest. In the supplementary material, we show how
to make similar calculations to determine the minimum time series
required to estimate overall population growth rate, the variability in
population size, and future extinction risk.

In the appendix, we explore other potential questions of
interest\ldots{}

\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-3-1.pdf}

\subsection{Results}\label{results}

\begin{itemize}
\tightlist
\item
  summary of findings
\item
  theoretical approach
\item
  example plot
\item
  empirical approach
\item
  plots: distribution of min time, what min time correlations with
\item
  walk through one empirical example (maybe in appendix)
\end{itemize}

We evaluated the question of what minimum time series is required to
answer a question of interest. More specifically we ask, what is the
minimum time series length required to determine, via linear regression,
the longterm trend in population abundance? The minimum time series
length required is that which had high enough statistical power. We
investigated this question through two approaches: simulation-based and
empirical analyses. In both cases, we found the minimum time series
length required depended strongly on the trend strength, population
variability, and autocorrelation.

\subsubsection{Simulation approach}\label{simulation-approach}

We constructed a general population model where the trend over time
could be a model parameter. We chose a simple model, but any other
population model could be used. The specific model choice should be
tailored to the population of interest, based on species ecology. We
then ran simulations of different lengths to determine the minimum time
series length required to achieve a certain level of power (which we set
to 0.8). We were also able to alter the parameter values (trend strength
and population variability) of the population model to see how this
altered \(T_{min}\).

In line with past work, we found the \(T_{min}\) increases (i.e.~more
time is required) with decreases in trend strength or autocorrelation
and with increases in variability (citation).

\subsubsection{Empirical test}\label{empirical-test}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Distribution of the minimum time required in order to detect a
significant trend (at the 0.05 level) in abundance given power of
0.8.\label{fig:fig:min_time_dist}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{duh\label{fig:correlates}}
\end{figure}

We examined a database of 868 separate population time series
representing 475 species. This database consists of species with a
variety of life history characteristics. We limited our analyses to
populations with at least 35 years of continuous sampling.

Across all the populations we examined, we found an average minimum time
series length required (\(T_{min}\)) of 16.6463134
(\(\sigma\)=8.6664727), with a wide distribution (Fig.
\ref{fig:min_time_dist}). As expected based on theoretical results (cite
papers), the minimum time series length required is strongly correlated
with trend strength, population variance, and autocorrelation (Fig
\ref{fig:correlates}). Using a generalized linear model, with a Poisson
error structure, all three of these explanatory variables were
significant and had large effect sizes (cite model here). Combined trend
strength, population variance, and autocorrelation account for about
70.7793122 of the variance in minimum time series length required (as
proposed by CITE Zuur).

We also examined a subset of populations where more explanatory
variables were known (e.g.~body mass, longevity, location). None (need
to do regression) of these explanatory variables were predictive of the
minimum time series length required. Some differences if \(t_{min}\) did
occur between biological classes, but this can be attributed to the
amount populations are know to fluctuate.

\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-6-1.pdf}

\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-7-1.pdf}

\subsection{Discussion}\label{discussion}

\begin{itemize}
\tightlist
\item
  Findings and how they relate to past work
\item
  Extensions to other systems. Useful to use similar species
\item
  Biological variables not good predictors, but we also used all
  long-lived vertebrate species
\item
  Other questions to ask and results in appendix
\item
  Future recommendations to others and future work
\item
  implications
\item
  conclusion paragraph
\end{itemize}

The IUCN has an 10 year/3 generation rule\ldots{} If we apply this rule
to the species in our database we see that approximately 50.5760369\% of
populations need a longer minimum time series than that required by
IUCN.

\subsubsection{Time series length recommendations from other
papers}\label{time-series-length-recommendations-from-other-papers}

\begin{itemize}
\tightlist
\item
  greater than 5 years (Morrison and Hik 2008)
\item
  more than 10-15 years (Rueda-Cediel et al. 2015)
\item
  general advice (variability, correlations) (Rhodes and Jonzen 2011)
\item
  more than 10 years (Gerber, DeMaster, and Kareiva 1999)
\item
  specific recommendations depend on amount of power and strength of
  trend (Bart 2004, Hatch 2003) (Seavy and Reynolds 2007)
\end{itemize}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\hypertarget{ref-Gerber1999}{}
Gerber, L R, D P DeMaster, and P M Kareiva. 1999. ``Gray whales and the
value of monitering data in implementing the U.S. endangered species
act.'' \emph{Conservation Biology} 13 (5): 1215--9.

\hypertarget{ref-Gerrodette1987}{}
Gerrodette, Tim. 1987. ``A power analysis for detecting trends.''
doi:\href{https://doi.org/10.2307/1939220}{10.2307/1939220}.

\hypertarget{ref-Johnson2015}{}
Johnson, Paul C D, Sarah J E Barry, Heather M. Ferguson, and Pie Müller.
2015. ``Power analysis for generalized linear mixed models in ecology
and evolution.'' \emph{Methods in Ecology and Evolution} 6 (2): 133--42.
doi:\href{https://doi.org/10.1111/2041-210X.12306}{10.1111/2041-210X.12306}.

\hypertarget{ref-Keith2015}{}
Keith, David, H. Resit Akçakaya, Stuart H.M. Butchart, Ben Collen,
Nicholas K. Dulvy, Elizabeth E. Holmes, Jeffrey A. Hutchings, et al.
2015. ``Temporal correlations in population trends: Conservation
implications from time-series analysis of diverse animal taxa.''
\emph{Biological Conservation} 192. Elsevier B.V.: 247--57.
doi:\href{https://doi.org/10.1016/j.biocon.2015.09.021}{10.1016/j.biocon.2015.09.021}.

\hypertarget{ref-Legg2006}{}
Legg, Colin J, and Laszlo Nagy. 2006. ``Why most conservation monitoring
is , but need not be , a waste of time.'' \emph{Journal of Environmental
Management} 78: 194--99.
doi:\href{https://doi.org/10.1016/j.jenvman.2005.04.016}{10.1016/j.jenvman.2005.04.016}.

\hypertarget{ref-Magurran2010}{}
Magurran, Anne E, Stephen R Baillie, Stephen T Buckland, Jan Mcp Dick,
David A Elston, E Marian Scott, Rognvald I Smith, Paul J Somerfield, and
Allan D Watt. 2010. ``Long-term datasets in biodiversity research and
monitoring : assessing change in ecological communities through time.''
\emph{Trends in Ecology and Evolution} 25: 574--82.
doi:\href{https://doi.org/10.1016/j.tree.2010.06.016}{10.1016/j.tree.2010.06.016}.

\hypertarget{ref-Morrison2008}{}
Morrison, Sharn, and David S. Hik. 2008. ``When? Where? and for how
long? Census design considerations for an Alpine Lagomorpg, the Collared
pika.'' In \emph{Lagomorph Biology}, 103--13. Springer Berlin
Heidelberg.
doi:\href{https://doi.org/10.1007/978-3-540-72446-9}{10.1007/978-3-540-72446-9}.

\hypertarget{ref-GPDD2010}{}
NERC Centre for Population Biology, Imperial College. 2010. ``The Global
Population Dynamics Database Version 2.''

\hypertarget{ref-Rhodes2011}{}
Rhodes, Jonathan R., and Niclas Jonzen. 2011. ``Monitoring temporal
trends in spatially structured populations: how should sampling effort
be allocated between space and time?'' \emph{Ecography} 34 (6): 1040--8.
doi:\href{https://doi.org/10.1111/j.1600-0587.2011.06370.x}{10.1111/j.1600-0587.2011.06370.x}.

\hypertarget{ref-Rueda-Cediel2015}{}
Rueda-Cediel, Pamela, Kurt E Anderson, Tracey J Regan, Janet Franklin,
and M Regan. 2015. ``Combined Influences of Model Choice , Data Quality
, and Data Quantity When Estimating Population Trends.'' \emph{PLoSONE}
10 (7): e0132255.
doi:\href{https://doi.org/10.1371/journal.pone.0132255}{10.1371/journal.pone.0132255}.

\hypertarget{ref-Seavy2007}{}
Seavy, Nathaniel E., and Michelle H. Reynolds. 2007. ``Is statistical
power to detect trends a good assessment of population monitoring?''
\emph{Biological Conservation} 140 (1-2): 187--91.
doi:\href{https://doi.org/10.1016/j.biocon.2007.08.007}{10.1016/j.biocon.2007.08.007}.

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=0.75in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{float} \renewcommand{\thepage}{S\arabic{page}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

@@

\vspace{2cm}

\begin{center}
 \textbf{Supplementary Material: Article title here}

Authors: Easton R. White$^{1*}$ , John D. Nagy$^{2,3}$, and Andrew T. Smith$^4$ 
\vspace{3 mm}

Address: \\ \emph{$^1$ Center for Population Biology, University of California-Davis, Davis, California 95616 USA; \\ $^2$ Department of Life Science, Scottsdale Community College, Scottsdale, Arizona 85253 USA; \\ $^3$ School of Mathematical and Statistical Sciences, Arizona State University, Tempe, Arizona 85287 USA; \\ $^4$ School of Life Sciences, Arizona State University, Tempe, Arizona 85287 USA}

*Corresponding author: eawhite@ucdavis.edu

Dec XX, 2016
 \end{center}

\vspace{2cm}

This supplementary material includes:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Mathematical description of our model
\item
  Inverse pattern-orientated estimation procedure
\item
  Effects of spatial structure and patch heterogeneity
\item
  Mainland-Island effect
\item
  Invasion from southern area of study site
\item
  Parameter sensitivity
\item
  Effect of unsampled patches
\item
  Long term dynamics
\item
  Individual patch dynamics over time
\end{enumerate}

Code for all the figures and tables can be found on
\href{https://github.com/erwhite1}{Github}.

\vspace{2cm}

\subsection{Example of subsamling and power
calculations}\label{example-of-subsamling-and-power-calculations}

Here, we illustrate in detail how we performed the subsampling and power
calculations for a specific population. As an example, we examine a
35-year time series of (). Simple linear regression indicates a
significant decline for this population with a blank and blank. We
assume that this significant decline over 35 years is in fact the ``true
trend``. In statistical jargon, the 35-year trend is an effect that is
actually present. We can then use this as a benchmark to see if
subsamples of the time series show a similar result.

We first extract all contiguous subsamples of the time series. This
leads to 34 two-year subsamples, 33 two-year subsamples, and so forth
until a single 35-year subsample. We can call each set of subsamples, of
the same length, a set. For each subsample, we conduct linear regression
and extract model coefficients and p-values. Then, the fraction of
subsamples within a set that show significant trends (signfiiant slope
coefficient) is the statistical power. It is important to note that we
only consider subsamples to be significant if they are significant in
the same direction as the complete 35-year time series.

We can then plot statistical power as a function of time series length.
As expected, we can see that power increases with the more years that
are sampled.

Then, we determine an appropriate level of statistical power that we
find acceptable. Traditionally, this has been at 0.8, however, this is
purely historical. Statisical power of 0.8 implies\ldots{}

With statistical power of 0.8, we then determine the minimum time series
length (\(t_min\)) required to achieve that level of statistical power.
Here, \(t_min\) is the first point as which all points to the right are
above 0.8.

\subsection{Results from other questions of
interest}\label{results-from-other-questions-of-interest}

\section{\texorpdfstring{Supplementary material
\ref{fig:growth_rate}}{Supplementary material }}\label{supplementary-material}

\begin{figure}[htbp]
\centering
\includegraphics{supp_mat_draft_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{duh\label{fig:growth_rate}}
\end{figure}

\bibliography{White_bib.bib}
